Increasing User Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Final version submitted

\begin{figure*}[htbp]
    \centering
        \includegraphics[width=5in]{img/login.png}
    \caption{ User interface: Landing page, and Login and registration dialogs.  
    }
    \label{fig:login}
\end{figure*}

@@

The gamification technique employed in this work is based on a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. In these case
a point is awarded for each PUT sent to the server. Levels are a long
term achievement, in this case the level depends on the score:
\[ \text{level}(\text{score};a,b)= 
    \begin{cases} 
      0,                                    &  \text{score}\leq a\\    
      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
      1,                                    & \text{score}\geq b 
   \end{cases}
\]
The function returns a normalized value that is multiplied by the maximum level 
in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of 
levels at the beginning.
% \begin{figure*}[htbp]
%     \centering
%         \includegraphics[width=2.5in]{img/s.png}
%     \caption{$S$ function used to assign the user's level.
%     }
%     \label{fig:s}
% \end{figure*}

@@

\subsection{Problem}
In the browser, each page visiting the experiment loads an WebWorker
that runs a local island of an evolutionary algorithm to solve a
multi-modal problem called {\em l-trap}, which has been used extensively 
as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}. 
This function counts the number of bits in a sequence $l$ and assigns
the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
bits, this makes the  the fitness fall into a {\em trap} 
as the number of bits is increased, decreasing linearly until a change in slope 
is reached at point $z$, adding deceptive component for evolutionary algorithms. 
To increase the difficulty trap functions can be concatenated. 
In our case we have used $40$ concatenated traps. The trap function  defined as:   
\[ f(u)= 
    \begin{cases} 
      \frac{a}{z}(z-u) & \text{if } u\leq z\\
      \frac{b}{l-z} (u-z)& \text{otherwise} 
   \end{cases}
\]
% \begin{figure*}[htbp]
%     \centering
%         \includegraphics[width=2.5in]{img/trap.png}
%     \caption{Trap function.
%     }
%     \label{fig:trap}
% \end{figure*}
%

@@

%
\begin{figure*}[htbp]
    \centering
        \includegraphics[width=4in]{img/puts_box.png}
    \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
     In the \emph{y} axis are number or PUTs in a logarithmic scale.
    }
    \label{fig:box}
\end{figure*}
%
The box plot \ref{fig:box} of the amount of participation between the two groups, highlights
the difference and also the number of outliers in the groups. Two registered 
outliers where very important participants maybe competing between them. But
there where more outliers between anonymous IPs, and some of them could be related
to the same users. The lower quartile in in the users box-plot has more
spread, indicating a participation similar to the median of anonymous IPs.

%
\begin{figure*}[htbp]
    \centering
        \includegraphics[width=5in]{img/workers_put_ip.png}
    \caption{ Participation of registered users. Each circle represents a user relating  
        in the \emph{x} axis the number of workers used and in the \emph{y} axis 
        number or PUTs in a logarithmic scale. The area of the circle is the number 
        of unique IPs used by the user.
    }
    \label{fig:worker-put-ips}
\end{figure*}
%
Figure~\ref{fig:worker-put-ips} gives an interesting view on the amount of
resources shared by registered users. The area of the circle is proportional
to the number of unique IPs used by each user, the user with more participation
used a total of 66 different IPs, this number could be related with the number
of devices used during the participation. There are some users that using less
devices participated more, this could mean a more powerful device, faster 
Internet connection or simply more time spent participating.

User engagement is related to the amount of time a user spends in the 
application. Figure~\ref{fig:puts-time} shows to important aspects of
engagement, the overall time spent and the amount of resources shared
in that time. Registered users where more engaged during the experiment.
They contribute more resources during longer periods of time.

%
\begin{figure*}[htbp]
    \centering
        \includegraphics[width=4in]{img/puts_by_time.png}
    \caption{Number of PUTs by User or IP in 15 minute slots.
    }
    \label{fig:puts-time}
\end{figure*}
%
\begin{figure*}[htbp]
    \centering
        \includegraphics[width=5in]{img/workers_best_user.png}
    \caption{ Number of workers used by the top ranked user in 15 minute slots.
    }
    \label{fig:top-user}
\end{figure*}
%
The participation of the top ranked user is presented in 15 seconds time slots in
Figure~\ref{fig:top-user}. In the beginning this user was using more than 30
workers, this means that more than 30 tabs of the page where open at the
same time, perhaps using several computing devices. Some users reported that 
they wanted to test the limits of their own systems, checking the percentage of
CPU they where using.

@@

>>>>

\begin{figure*}[htb]
    \centering
        \includegraphics[width=5in]{img/login.png}
    \caption{ User interface: Landing page, and Login and registration dialogs.  
    }
    \label{fig:login}
\end{figure*}

@@

The gamification technique employed in this work is based on a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. In these case
a point is awarded for each PUT sent to the server. Levels are a long
term achievement, in this case the level depends on the score:
\[ \text{level}(\text{score};a,b)= 
    \begin{cases} 
      0,                                    &  \text{score}\leq a\\    
      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
      1,                                    & \text{score}\geq b 
   \end{cases}
\]
The function returns a normalized value that is multiplied by the maximum level 
in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of 
levels at the beginning.
% \begin{figure*}[htb]
%     \centering
%         \includegraphics[width=2.5in]{img/s.png}
%     \caption{$S$ function used to assign the user's level.
%     }
%     \label{fig:s}
% \end{figure*}

@@

\subsection{Problem}
In the browser, each page visiting the experiment loads an WebWorker
that runs a local island of an evolutionary algorithm to solve a
multi-modal problem called {\em l-trap}, which has been used extensively 
as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}. 
This function counts the number of bits in a sequence $l$ and assigns
the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
bits, this makes the  the fitness fall into a {\em trap} 
as the number of bits is increased, decreasing linearly until a change in slope 
is reached at point $z$, adding deceptive component for evolutionary algorithms. 
To increase the difficulty trap functions can be concatenated. 
In our case we have used $40$ concatenated traps. The trap function  defined as:   
\[ f(u)= 
    \begin{cases} 
      \frac{a}{z}(z-u) & \text{if } u\leq z\\
      \frac{b}{l-z} (u-z)& \text{otherwise} 
   \end{cases}
\]
% \begin{figure*}[htb]
%     \centering
%         \includegraphics[width=2.5in]{img/trap.png}
%     \caption{Trap function.
%     }
%     \label{fig:trap}
% \end{figure*}
%

@@

%
\begin{figure*}[htb]
    \centering
        \includegraphics[width=4in]{img/puts_box.png}
    \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
     In the \emph{y} axis are number or PUTs in a logarithmic scale.
    }
    \label{fig:box}
\end{figure*}
%
The box plot featured in Figure \ref{fig:box} shows the amount of participation between the two groups, highlights
the difference and also the number of outliers in the groups. Two registered 
outliers where very important participants maybe competing between them. But
there where more outliers between anonymous IPs, and some of them could be related
to the same users. The lower quartile in in the users box-plot has more
spread, indicating a participation similar to the median of anonymous IPs.

%
\begin{figure*}[htb]
    \centering
        \includegraphics[width=5in]{img/workers_put_ip.png}
    \caption{ Participation of registered users. Each circle represents a user relating  
        in the \emph{x} axis the number of workers used and in the \emph{y} axis 
        number or PUTs in a logarithmic scale. The area of the circle is the number 
        of unique IPs used by the user.
    }
    \label{fig:worker-put-ips}
\end{figure*}
%
Figure~\ref{fig:worker-put-ips} gives an interesting view on the amount of
resources shared by registered users. The area of the circle is proportional
to the number of unique IPs used by each user, the user with more participation
used a total of 66 different IPs, this number could be related with the number
of devices used during the participation. There are some users that using less
devices participated more, this could mean a more powerful device, faster 
Internet connection or simply more time spent participating.

\begin{figure*}[htb]
    \centering
        \includegraphics[width=4in]{img/puts_by_time.png}
    \caption{Number of PUTs by User or IP in 15 minute slots.
    }
    \label{fig:puts-time}
\end{figure*}
%
User engagement is related to the amount of time a user spends in the 
application. Figure~\ref{fig:puts-time} shows to important aspects of
engagement, the overall time spent and the amount of resources shared
in that time. Registered users where more engaged during the experiment.
They contribute more resources during longer periods of time.

\begin{figure*}[htb]
    \centering
        \includegraphics[width=5in]{img/workers_best_user.png}
    \caption{ Number of workers used by the top ranked user in 15 minute slots.
    }
    \label{fig:top-user}
\end{figure*}
%
The participation of the top ranked user is presented in 15 seconds time slots in
Figure~\ref{fig:top-user}. In the beginning this user was using more than 30
workers, this means that more than 30 tabs of the page where open at the
same time, perhaps using several computing devices. Some users reported that 
they wanted to test the limits of their own systems, checking the percentage of
CPU they where using.

@@

The fact that gamification actually work highlights the social nature of volunteer computing systems, piling on the fact that successful {\em social clouds} must embed games and other social mechanisms to enhance participation. That is why future lines of work will look for intermediate ways of including
gamification without registration, as well as some ways of making the user increase control over the algorithm that would go beyond simply staying for longer or leaving the page. Since it is a sociotechnical system, including social features in it and studying emerging social networks might be an interesting line of work too.

Other factors, such as the language used, the way the experiment is announced, are also very important in this social context, which is why they will be included in future lines of work too.
